---
title: "THESIS"
author: "stina"
date: "`r Sys.Date()`"
output: html_document
---
```{r, warning=FALSE, message=FALSE}
library(readxl)
library(ggplot2)
library(tidyverse)
library(knitr)
library(gridExtra)

# Read the Excel file
iligan_data <- readxl::read_xlsx(path="C:\\Users\\ACER\\Downloads\\ThesisIT!\\IC\\Iligan_Dengue_Cases.xlsx", sheet=1) 
```

```{r}
iligan_data$Week <- as.numeric(gsub("W", "", iligan_data$WEEK_NUM))

iligan_data$Date <- as.Date(paste(iligan_data$YEAR, iligan_data$Week, 1, sep = "-"), format = "%Y-%U-%u")
```

```{r, warning=FALSE, fig.align = 'center',fig.cap = "Data 1: Weekly Dengue Cases in Iligan City (2010–2024)"}
ggplot(iligan_data, aes(x = Date, y = ILIGAN)) +
  geom_line(color = "black", size = 1) +
  labs(title = "Data 1: Weekly Dengue Cases in Iligan City (2010–2024)",
       x = "Time",
       y = "Y_t") +
  theme_minimal() +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y") 

```



```{r, warning=FALSE, fig.align = 'center',fig.cap = "Data 2: Weekly Dengue Cases in Tubod, Iligan City (2010–2024)"}
fig2 <-ggplot(iligan_data, aes(x = Date, y = TUBOD)) +
  geom_line(color = "black", size = 1) +
  labs(title = "Data 2: Weekly Dengue Cases in Tubod, Iligan City (2010–2024)",
       x = "Time",
       y = "Y_t") +
  theme_minimal() +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y") 

```



```{r, warning=FALSE, fig.align = 'center',fig.cap = "Data 3: Weekly Dengue Cases in Tibanga, Iligan City (2010–2024)"}
fig3 <-ggplot(iligan_data, aes(x = Date, y = TIBANGA)) +
  geom_line(color = "black", size = 1) +
  labs(title = "Data 3: Weekly Dengue Cases in Tibanga, Iligan City (2010–2024)",
       x = "Time",
       y = "Y_t") +
  theme_minimal() +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y") + 
  scale_y_continuous(breaks = seq(0, 25, by=2)) 

grid.arrange(fig2,fig3)
```




```{r, warning=FALSE, fig.align = 'center',fig.cap = "Data 4: Weekly Dengue Cases in Suarez, Iligan City (2010–2024)"}
fig4 <-ggplot(iligan_data, aes(x = Date, y = SUAREZ)) +
  geom_line(color = "black", size = 1) +
  labs(title = "Data 4: Weekly Dengue Cases in Suarez, Iligan City (2010–2024)",
       x = "Time",
       y = "Y_t") +
  theme_minimal() +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y") 

```



```{r, warning=FALSE, fig.align = 'center',fig.cap ="Data 5: Weekly Dengue Cases in Buruun, Iligan City(2010–2024)"}
fig5 <- ggplot(iligan_data, aes(x = Date, y = BURUUN)) +
  geom_line(color = "black", size = 1) +
  labs(title = "Data 5: Weekly Dengue Cases in Buruun, Iligan City(2010–2024)",
       x = "Time",
       y = "Y_t") +
  theme_minimal() +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y") +  
  scale_y_continuous(breaks = seq(0, 25, by=2))  

grid.arrange(fig4,fig5)
```


```{r, warning=FALSE, }
ggplot(iligan_data, aes(x = Date, y = rainfall)) +
  geom_line(color = "black", size = 1) +
  labs(title = "Weekly cumulative rainfall in Iligan City",
       x = "Time",
       y = "Y_t") +
  theme_minimal() +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y") 

```



```{r, warning=FALSE, }
ggplot(iligan_data, aes(x = Date, y = temperature)) +
  geom_line(color = "black", size = 1) +
  labs(title = "Weekly cumulative rainfall in Iligan City",
       x = "Time",
       y = "Y_t") +
  theme_minimal() +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y") 

```


```{r}
# Descriptive Statistics for dengue data
get_descriptive_stats <- function(df, column_name) {
  df %>%
    summarise(
      mean = mean(get(column_name), na.rm = TRUE),
      variance = var(get(column_name), na.rm = TRUE),
      min = min(get(column_name), na.rm = TRUE),
      Q1 = quantile(get(column_name), 0.25, na.rm = TRUE),
      median = median(get(column_name), na.rm = TRUE),
      Q3 = quantile(get(column_name), 0.75, na.rm = TRUE),
      max = max(get(column_name), na.rm = TRUE),
      zero_count = sum(get(column_name) == 0, na.rm = TRUE),
      zero_percentage = sum(get(column_name) == 0, na.rm = TRUE) / length(get(column_name)) * 100  
    )
}

# List of column names for each city
cities <- c("ILIGAN",  "TUBOD", "TIBANGA", "SUAREZ", "BURUUN" )


descriptive_stats <- lapply(1:length(cities), function(i) {
  city <- cities[i]
  stats <- get_descriptive_stats(iligan_data, city)
  stats$data <- paste("Data", i)  
  stats$location <- city
  return(stats)
})

# Combining results into a single data frame
descriptive_stats_df <- do.call(rbind, descriptive_stats)
descriptive_stats_df <- descriptive_stats_df[, c("data", "location", "mean", "variance", "min", "Q1", "median", "Q3", "max", "zero_percentage")]

# Display the summary table
print(descriptive_stats_df)

```

```{r}
# Descriptive Statistics for the meteorological data (rainfall and temperature)
cov_descriptive_stats <- function(df, column_name) {
  df %>%
    summarise(
      mean = mean(get(column_name), na.rm = TRUE),
      median = median(get(column_name), na.rm = TRUE),
      min = min(get(column_name), na.rm = TRUE),
      max = max(get(column_name), na.rm = TRUE)
    )
}

# List of exogenous variables (covariates)
exogenous <- c("weekly_rainfall", "weekly_temp_max")

# Apply the descriptive statistics function to each exogenous variable
cov_ds <- lapply(exogenous, function(variable) {
  cov_stats <- cov_descriptive_stats(iligan_data, variable)
  cov_stats$exogenous_variable <- ifelse(variable == "weekly_rainfall", "Rainfall (mm)", "Temperature (°C)")  # Add labels
  return(cov_stats)
})

# Combine the results into a single data frame
cov_descriptive_stats_df <- do.call(rbind, cov_ds)
cov_descriptive_stats_df <- cov_descriptive_stats_df[, c("exogenous_variable", "mean", "median", "min", "max")]

# Display the summary table
print(cov_descriptive_stats_df)

```




#-------------------------------------------------
#Note: Run all of the code chunks above before running this
# Run this sa console for the zero-inflation Index

calculate_zero_inflation <- function(count_data) {
  # Remove NA
  count_data <- na.omit(count_data)
  
  lambda <- mean(count_data)
  sigma2 <- var(count_data)
  p0 <- mean(count_data == 0)
  
  # zi_P
  zi_P <- 1 + log(p0) / lambda
  
  
  return(data.frame(Mean = lambda, Variance = sigma2, p0 = p0, zi_P = zi_P))
}

# List of barangays
barangays <- c("ILIGAN", "TUBOD", "TIBANGA", "SUAREZ", "BURUUN")

# Apply function to each column
zi_results <- lapply(barangays, function(area) {
  result <- calculate_zero_inflation(iligan_data[[area]])
  result$Area <- area
  return(result)
})

# Combine and format
zi_results_df <- do.call(rbind, zi_results) %>%
  select(Area, everything()) %>%
  mutate(across(where(is.numeric), round, 4))

# Display the table
kable(zi_results_df, caption = "Zero-Inflation Indices for Weekly Dengue Cases")



